---
title: "Project 2 Data Summaries"
output:
   html_document:
    self_contained: true
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = paste0(substr(inputFile,1,nchar(inputFile)-4),Sys.Date(),'.html')) })
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)

packages <- c("DBI", "dplyr","kableExtra","table1","consort")

for (pkg in packages) {
  # If package is not installed then install it. 
  if(!(pkg %in% installed.packages()[,1])){install.packages(pkg)}
  
  # Load package. 
  library(pkg,character.only = TRUE)
}

```

```{r set file locations}
# Set file locations ---------

if (Sys.info()["sysname"] == "Darwin") {
  setwd("~/Library/CloudStorage/GoogleDrive-joelparkerbiostatistics@gmail.com/My Drive/Computer/Bio5/PAN Reports/BaselineTabulations")
} else if (Sys.info()["sysname"] == "Linux") {
  setwd("/data/rscripts/BaselineTabulations/")
}

library(DBI) # Connecting to database
library(tidyverse)

# load("dbconnect_info.Rdata")

con <- dbConnect(
   RPostgres::Postgres(),
   host = "bio5-pandb-nlb-83001662f7d2c9f3.elb.us-west-2.amazonaws.com",
   user = "pan_investigator",
   password = Sys.getenv("password"),
   dbname = "healthy_minds_for_life",
   sslmode = "require",
   sslrootcert = "pandb_root.crt"
 )


```



```{r wrangle data sets}

# # Load tables

demo <- dbReadTable(con, "p2_redcap_demographics") %>%
  select(redcap_id, hml_id, hml_id_created_date, api_import_date,
         age, decade, sex, race, hispanic_latino, group_id_number, memory_rank)  %>%
  # Create variables for app table/plots
  mutate(age_group = case_when(age %in% 50:59 ~ "50-59",
                               age %in% 60:69 ~ "60-69",
                               age %in% 70:79 ~ "70-79",
                               TRUE ~ as.character(age)),
         race_group = case_when(hispanic_latino == "Yes" ~ "Hispanic/Latino",
                                race == "White" ~ "Non-Hispanic White",
                                race == "Black or African American" ~ "Non-Hispanic Black",
                                TRUE ~ "Other"),
         memory_rank = stringr::str_to_upper(memory_rank),
         area = case_when(group_id_number == 4669 ~ "Atlanta",
                          group_id_number == 4668 ~ "Baltimore",
                          group_id_number == 4667 ~ "Miami",
                          group_id_number == 4670 ~ "Tucson"))

elig <- dbReadTable(con, "p2_redcap_eligibility") %>%
  select(redcap_id, hml_id, not_eligible, hif_eligible_behav, hif_eligible_mri, hif_eligible_bd)

cons <- dbReadTable(con, "p2_redcap_consent_form") %>%
  select(redcap_id, hml_id, main_consent_date, main_consent_upload)

phq9 <- dbReadTable(con, "p2_redcap_cognitive_function_phq9")

moca <- dbReadTable(con, "p2_redcap_cognitive_function_moca")

naart <- dbReadTable(con, "p2_redcap_cognitive_function_naart")

avlt <- dbReadTable(con, "p2_redcap_cognitive_function_avlt")

biometrics <- dbReadTable(con, "p2_redcap_biometrics")

# dm_checklist <- dbReadTable(con, "p2_redcap_dm_checklist") %>%
  # select(hml_id, dm_checklist_complete)

base <- full_join(phq9, moca, by = c("hml_id", "redcap_id", "mindcrowd_id")) %>%
  full_join(naart, by = c("hml_id", "redcap_id", "mindcrowd_id")) %>%
  full_join(avlt, by = c("hml_id", "redcap_id", "mindcrowd_id")) %>%
  full_join(biometrics, by = c("hml_id", "redcap_id", "mindcrowd_id")) %>%
  # full_join(neuropsych, by = c("hml_id", "redcap_id", "mindcrowd_id")) %>%
  # One of the redcap_ids is missing a hml_id
  left_join(select(demo, redcap_id, hml_id) %>% rename(new_hml_id = hml_id), by = "redcap_id") %>%
  mutate(hml_id = ifelse(is.na(hml_id), new_hml_id, hml_id)) %>%
  select(-new_hml_id)

# surv <- dbReadTable(con, "p2_redcap_survey_info") %>%
#   select(redcap_id, hml_id, hml_survey_tracker_15_complete)

```

`r if(sum(duplicated(demo$redcap_id)) > 0){"dm_redcapx_demographic update contains duplicated rows"}`
`r if(sum(duplicated(elig$redcap_id)) > 0){"dm_redcapx_eligibility update contains duplicated rows"}`
`r if(sum(duplicated(cons$redcap_id)) > 0){"dm_redcapx_consents update contains duplicated rows"}`
`r if(sum(duplicated(base$redcap_id)) > 0){"dm_redcapx_baseline_assessment update contains duplicated rows"}`

The following tables summarize data collected in REDCap for PAN Project 2.

Report Date: **`r format(Sys.Date(), "%b %d, %Y")`**

#### Participants with HML IDs

These participants are in the MC Demographics form and have been assigned HML IDs.

```{r recruit summary}

participants_dat <- demo %>%
  left_join(cons, by = "redcap_id") %>%
  mutate(consent_uploaded = ifelse(!is.na(main_consent_upload), "Consented", "Non-Consented"))

label(participants_dat$sex) <- "Sex"
label(participants_dat$age_group) <- "Age"
label(participants_dat$race_group) <- "Race"
label(participants_dat$consent_uploaded) <- "Consent Uploaded?"

table1(~sex + age_group + race_group + consent_uploaded | area, data = participants_dat)

```

#### Consented Participants

The following table contains only participants whose consent has been uploaded to REDCap. The rest of this report follows these participants.

```{r consented summary}

table1(~sex + age_group + race_group | area, 
       data = filter(participants_dat,  !is.na(main_consent_upload)))

```

```{r nonconsented}

non_consents <- cons %>%
  filter(redcap_id %in% base$redcap_id & is.na(main_consent_upload)) %>%
  left_join(select(demo, redcap_id, group_id_number), by = "redcap_id") %>%
  mutate(area = case_when(group_id_number == 4669 ~ "Atlanta",
                          group_id_number == 4668 ~ "Baltimore",
                          group_id_number == 4667 ~ "Miami",
                          group_id_number == 4670 ~ "Tucson")) %>%
  select(area, redcap_id, hml_id, main_consent_date, main_consent_upload)

```

<br>

There are `r nrow(non_consents)` participants without uploaded consents, but still have data collected in other REDCap forms. `r nrow(non_consents[!is.na(non_consents$main_consent_date),])` of these have a "Main Consent Date" in the Consent form:

```{r nonconsented table}
if(nrow(non_consents) > 0){
  non_consents %>%
    mutate(across(-c(area, redcap_id, hml_id), ~ifelse(is.na(.), "Missing", .))) %>%
    kable(col.names = c("Area", "Record ID", "HML ID", "Main Consent Date", "Main Consent Uploaded"),
          align = "c", escape = FALSE) %>%
    column_spec(1:2, bold = TRUE) %>%
    column_spec(1, extra_css = "border-bottom: 1px dashed; background: white;") %>%
    row_spec(0, bold = TRUE) %>%
    collapse_rows(columns = 1) %>%
    kable_classic(full_width = FALSE, lightable_options = "striped", position = "left")
}else{
  print("All participants with REDCap data have uploaded consents.")
}


```

I have removed those participants from the rest of this report.

------

```{r filtering consented participants}

base <- filter(base,
               redcap_id %in% participants_dat$redcap_id[participants_dat$consent_uploaded == "Consented"])

elig <- filter(elig,
               redcap_id %in% participants_dat$redcap_id[participants_dat$consent_uploaded == "Consented"])

```

### Cognitive Testing and Biometrics

This table summarizes the cognitive testing in Scored Forms and the Biometrics Survey Status from the Experimental Tasks form. Participants without any missing form data are marked as "Non-Missing", if they have at least one missing entry then they are marked as "Missing."

```{r cognitive and biometrics summary}

assess <- base %>%
  # Add in demo data
  full_join(select(demo, redcap_id, area, hml_id,
                   age_group, race_group, memory_rank, sex), by = c("redcap_id", "hml_id")) %>%
  # Filter for consented participants
  left_join(cons, by = c("redcap_id", "hml_id")) %>%
  filter(!is.na(main_consent_upload)) %>%
  # Get postgres update date
  mutate(update_date = format(Sys.Date(), "%y%m%d")) %>%
  # Cognitive Summaries ~~~~~~~~~~~~~~
  mutate(across(phq9_endorse_q9:avlt_trial_a7_raw, ~ifelse(is.na(.), 1, 0))) %>%
  mutate("PHQ-9" = ifelse((phq9_endorse_q9 + phq9_total) > 0, "Missing", "Non-Missing"),
         # 20-Feb-24: Removed the education bonus and mis_total variables since not used in final total:
         MoCA = ifelse(select(., moca_abstraction, starts_with("moca_attention"), moca_delay_recall, starts_with("moca_language_"),
                              moca_visuospatial, moca_naming, moca_orientation, moca_total) %>% rowSums() > 0, "Missing", "Non-Missing"),
         NAART = ifelse(select(., naart_total_correct) %>% rowSums() > 0, "Missing", "Non-Missing"),
         AVLT = ifelse(select(., starts_with("avlt_") & ends_with("_raw")) %>% rowSums() > 0, "Missing", "Non-Missing")) %>%
  # Biometrics Form ~~~~~~~~~~~~~~~~~~
  mutate(across(bp_r_systolic_1:waist_hip_ratio, ~ifelse(is.na(.), 1, 0))) %>%
  mutate("Biometrics Form" = ifelse(select(., bp_r_systolic_1:waist_hip_ratio) %>% rowSums() > 0, "Missing", "Non-Missing")) %>%
  # All table variables
  mutate(across(c(`PHQ-9`, `MoCA`, `NAART`, `AVLT`, 
                  # `Neuropsychology Upload`, 
                  # `PHQ-9 Upload`, 
                  `Biometrics Form`), 
                ~factor(., levels = c("Non-Missing", "Missing")))) %>%
  # Get overall missingness:
  mutate(Overall = factor(ifelse(`PHQ-9` == "Non-Missing" & MoCA == "Non-Missing" & NAART == "Non-Missing" & 
                                   AVLT == "Non-Missing" & 
                                   # `Neuropsychology Upload` == "Non-Missing" & 
                                   # `PHQ-9 Upload` == "Non-Missing" & 
                                   `Biometrics Form` == "Non-Missing", "Complete Neuropsych and Biometrics", 
                                 "Incomplete Neuropsych and Biometrics"),
                          levels = c("Complete Neuropsych and Biometrics", "Incomplete Neuropsych and Biometrics"))) %>%
  select(redcap_id, hml_id = hml_id, area, update_date, age_group, race_group, memory_rank, sex, main_consent_date, ends_with("_complete"),
         main_consent_upload, `PHQ-9`, `MoCA`, `NAART`, `AVLT`, 
         # `Neuropsychology Upload`, 
         # `PHQ-9 Upload`, 
         `Biometrics Form`,
         Overall) %>%
  # Add in whether participant is eligible for study
  left_join(elig, by = c("redcap_id", "hml_id" = "hml_id")) %>%
  mutate(not_eligible = case_when(not_eligible == 0 ~"Not Eligible", 
                                  !is.na(hif_eligible_behav) ~ "Eligible"))

table1(~`PHQ-9` + MoCA + NAART + AVLT  +  `Biometrics Form` + Overall| area, 
       data = assess)
  
```

```{r list of hmls with complete data, eval = FALSE}

hml_complete <- assess %>%
  filter(`PHQ-9` == "Non-Missing", MoCA == "Non-Missing", NAART == "Non-Missing", AVLT == "Non-Missing", `Neuropsychology Upload` == "Non-Missing", `PHQ-9 Upload` == "Non-Missing", `Biometrics Form` == "Non-Missing") %>%
  select(area, hml_id)

atlanta_ignore_moca <- assess %>%
  filter(area %in% "Atlanta", `PHQ-9` == "Non-Missing", NAART == "Non-Missing", AVLT == "Non-Missing", `Neuropsychology Upload` == "Non-Missing", `PHQ-9 Upload` == "Non-Missing", `Biometrics Form` == "Non-Missing") %>%
  select(area, hml_id) %>%
  slice_sample(n = 20)

combined_hml_complete <- hml_complete %>%
  rbind(atlanta_ignore_moca) %>%
  arrange(desc(area), hml_id)

# write.csv(combined_hml_complete, file = "complete_data_hml_ids_240124.csv")

```

-----

<!-- ### DM Checklist and Form Completed Variables -->

<!-- This section summarizes the checklist found in "DM Checklist" which is meant to track the data entry/uploads and QC checks of the REDCap forms. -->

<!-- #### DM checklist status:  -->

<!-- This table summarizes the status (completed/unverified/incomplete) of the DM checklist form. -->

<!-- ```{r tabulate completion, results='asis'} -->

<!-- # Name factor levels -->
<!-- dm_checklist$dm_checklist_complete2 <- factor(case_when(dm_checklist$dm_checklist_complete == 0 ~ "Incomplete", -->
<!--                                          dm_checklist$dm_checklist_complete == 1 ~ "Unverified", -->
<!--                                          dm_checklist$dm_checklist_complete == 2 ~ "Complete"), -->
<!--                                       levels = c("Incomplete", "Unverified", "Complete")) -->

<!-- # Prettify names for table -->
<!-- dm_checklist_tab <- left_join(dm_checklist, select(demo, hml_id, area), by = "hml_id") %>% -->
<!--   filter(!is.na(hml_id)) -->
<!-- label(dm_checklist_tab$dm_checklist_complete2) <- "DM Checklist Status" -->

<!-- # Create table -->
<!-- table1(~ dm_checklist_complete2 | area, data = dm_checklist_tab) -->

<!-- ``` -->

<!-- #### User Form Completion vs Missing Data -->

<!-- We noticed some discrepancies between some forms that are marked as "Completed" but have missing entries. The following tables compare the forms' variables marked as complete--the table columns with labels ending in "_complete"--with what we have determined to be non-missing data. -->

<!-- ```{r dm to neuropsych compare} -->

<!-- # In cognitive and biometrics summary, assigned 0 to Non-missing entries except for biometrics form! -->
<!-- tests <- assess %>% -->
<!--   select(redcap_id, area, "PHQ-9":"Biometrics Form", ends_with("_complete"), biometrics_complete) %>% -->
<!--   mutate(across(ends_with("_complete"), ~factor(ifelse(. == 0, "Non-Missing", "Missing"), -->
<!--                                                 levels = c("Non-Missing", "Missing"))), -->
<!--          biometrics_complete = ifelse(biometrics_complete == "Non-Missing", "Missing", "Non-Missing")) -->

<!-- ``` -->

<!-- PHQ-9 -->

<!-- ```{r} -->
<!-- with(tests, addmargins(table(`PHQ-9`, `phq9_complete`))) -->
<!-- ``` -->


<!-- MoCA -->

<!-- ```{r} -->
<!-- with(tests, addmargins(table(`MoCA`, `moca_complete`))) -->
<!-- ``` -->


<!-- NAART -->

<!-- ```{r} -->
<!-- with(tests, addmargins(table(`NAART`, `naart_complete`))) -->
<!-- ``` -->


<!-- AVLT -->

<!-- ```{r} -->
<!-- with(tests, addmargins(table(`AVLT`, `avlt_complete`))) -->
<!-- ``` -->


<!-- Biometrics -->

<!-- ```{r} -->
<!-- with(tests, addmargins(table(`Biometrics Form`, `biometrics_complete`))) -->
<!-- ``` -->

#### Participants marked as ineligible or dropped from the study:

The following were ineligible for behavioral testing:

- `r elig[which(elig$hif_eligible_behav == "No"), "hml_id"]`

The following were ineligible for MRI scanning:

- `r elig[which(elig$hif_eligible_mri == "No"), "hml_id"]`

The following were ineligible for blood draw:

- `r elig[which(elig$hif_eligible_bd == "No"), "hml_id"]`

IDs dropped from the study:

```{r eligibility}

elig %>% 
  left_join(select(demo, redcap_id, area), by = "redcap_id") %>%
  filter(not_eligible == 0) %>%
  select(hml_id, redcap_id, area, hif_eligible_behav, hif_eligible_mri, hif_eligible_bd) %>%
  kable(col.names = c("HML ID", "REDCap ID", "Area", "Eligible for Behavior?", "Eligible for MRI?", "Eligible for Blood Draw?")) %>% 
  kable_classic(full_width = FALSE)

```

#### Data Flowchart

```{r flowchart data setup}

flow_dat <- demo %>%
  # Add consent data
  left_join(cons, by = c("redcap_id", "hml_id")) %>%
  mutate(consent_uploaded = ifelse(!is.na(main_consent_upload), "Consented", "Non-Consented")) %>%
  # Add eligibility data
  left_join(elig, by = c("redcap_id", "hml_id")) %>%
  # Add cognitive and biometrics data
  left_join(base, by = c("redcap_id", "hml_id")) %>%
  # Cognitive Summaries ~~~~~~~~~~~~~~
  mutate(across(phq9_endorse_q9:avlt_trial_a7_raw, ~ifelse(is.na(.), 1, 0))) %>%
  mutate("PHQ-9" = ifelse((phq9_endorse_q9 + phq9_total) > 0, "Missing", "Non-Missing"),
         # 20-Feb-24: Removed the education bonus and mis_total variables since not used in final total:
         MoCA = ifelse(select(., moca_abstraction, starts_with("moca_attention"), moca_delay_recall, starts_with("moca_language_"),
                              moca_visuospatial, moca_naming, moca_orientation, moca_total) %>% rowSums() > 0, "Missing", "Non-Missing"),
         NAART = ifelse(select(., naart_total_correct) %>% rowSums() > 0, "Missing", "Non-Missing"),
         AVLT = ifelse(select(., starts_with("avlt_") & ends_with("_raw")) %>% rowSums() > 0, "Missing", "Non-Missing")) %>%
  # Biometrics Form ~~~~~~~~~~~~~~~~~~
  mutate(across(bp_r_systolic_1:waist_hip_ratio, ~ifelse(is.na(.), 1, 0))) %>%
  mutate("Biometrics Form" = ifelse(select(., bp_r_systolic_1:waist_hip_ratio) %>% rowSums() > 0, "Missing", "Non-Missing")) %>%
  # All table variables
  mutate(across(c(`PHQ-9`, `MoCA`, `NAART`, `AVLT`, 
                  # `Neuropsychology Upload`, 
                  # `PHQ-9 Upload`, `Biometrics Form`
                  ), 
                ~factor(., levels = c("Non-Missing", "Missing")))) %>%
  # Get overall missingness:
  mutate(Overall = factor(ifelse(`PHQ-9` == "Non-Missing" & MoCA == "Non-Missing" & NAART == "Non-Missing" & 
                                   AVLT == "Non-Missing" & 
                                   # `Neuropsychology Upload` == "Non-Missing" & 
                                   # `PHQ-9 Upload` == "Non-Missing" & 
                                   `Biometrics Form` == "Non-Missing", "Complete Neuropsych and Biometrics", 
                                 "Incomplete Neuropsych and Biometrics"),
                          levels = c("Complete Neuropsych and Biometrics", "Incomplete Neuropsych and Biometrics")))

```

```{r flowchart}

# Flowchart outline

# Assign exclusion (remove) variables and reasoning if excluded; should be NA if participant is not excluded
# Assign inclusion variables that entered the next phase

# all hml_ids - redcap_id

# remove non-consented - is.na(main_consent_upload)
flow_dat$non_consented <- ifelse(!is.na(flow_dat$main_consent_upload), NA, "Missing Consent")
# consented
flow_dat$consented_group <- ifelse(!is.na(flow_dat$non_consented), NA, flow_dat$redcap_id)

# remove non-eligble
flow_dat$non_eligible <- case_when(is.na(flow_dat$not_eligible) ~ NA,
                                           flow_dat$hif_eligible_behav == "No" ~ "Ineligble for Behavioral Testing",
                                           flow_dat$hif_eligible_mri == "No" ~ "Ineligble for MRI",
                                           flow_dat$hif_eligible_bd == "No" ~ "Ineligble for Blood Draw")
# eligble participants
flow_dat$eligible <- ifelse(!is.na(flow_dat$non_consented) | !is.na(flow_dat$not_eligible), NA, flow_dat$redcap_id)

# remove missing cognitive/biometrics data
flow_dat$missing_data <- case_when(flow_dat$`PHQ-9` == "Missing" ~ "PHQ-9",
                                          flow_dat$MoCA == "Missing" ~ "MoCA",
                                          flow_dat$NAART == "Missing" ~ "NAART",
                                          flow_dat$AVLT == "Missing" ~ "AVLT",
                                          # flow_dat$`Neuropsychology Upload` == "Missing" ~ "Neuropsych Upload",
                                          flow_dat$`Biometrics Form` == "Missing"~ "Biometrics",
                                          TRUE ~ NA)

# non-missing participants
flow_dat$non_missing <- ifelse((!is.na(flow_dat$non_consented) | !is.na(flow_dat$not_eligible) |
                                         !is.na(flow_dat$missing_data)), NA, flow_dat$redcap_id)

consort_plot(flow_dat,
             orders = c(redcap_id = "Screened",
                        # area = "Site",
                        non_consented = "Missing Consent",
                        consented_group = "Consented",
                        non_eligible = "Ineligible",
                        eligible = "Eligible",
                        missing_data = "Missing Cognitive/Biometrics Data",
                        non_missing = "Non-Missing Data"),
             side_box = c("non_consented", "non_eligible", "missing_data")
             # ,allocation = "area"
             )

```

