---
title: "Project 2 Data Summaries"
output:
   html_document:
    self_contained: true
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = paste0(substr(inputFile,1,nchar(inputFile)-4),Sys.Date(),'.html')) })
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)

packages <- c("DBI", "dplyr","kableExtra","table1","consort")

for (pkg in packages) {
  # If package is not installed then install it. 
  if(!(pkg %in% installed.packages()[,1])){install.packages(pkg)}
  
  # Load package. 
  library(pkg,character.only = TRUE)
}

```

```{r set file locations}
# Set file locations ---------

if (Sys.info()["sysname"] == "Darwin") {
  setwd("~/Library/CloudStorage/GoogleDrive-joelparkerbiostatistics@gmail.com/My Drive/Computer/Bio5/PAN Reports/BaselineTabulations")
} else if (Sys.info()["sysname"] == "Linux") {
  setwd("/data/rscripts/BaselineTabulations/")
}

library(DBI) # Connecting to database
library(tidyverse)

con <- dbConnect(
   RPostgres::Postgres(),
   host = Sys.getenv("host"),
   user = Sys.getenv("user"),
   password = Sys.getenv("password"),
   dbname = Sys.getenv("dbname"),
   sslmode = Sys.getenv("sslmode"),
   sslrootcert = "rds-ca-2019-root (2).pem"
 )


```

```{r wrangle data sets}

# # Load tables

demo <- dbReadTable(con, "p2_redcap_demographics") %>%
  select(record_id, hml_id, hml_id_created_date,
         age, decade, sex, race, hispanic_latino, area = site, memory_rank)  %>%
  # Create variables for app table/plots
  mutate(age_group = case_when(age %in% 40:49 ~ "40-49",
                               age %in% 50:59 ~ "50-59",
                               age %in% 60:69 ~ "60-69",
                               age %in% 70:79 ~ "70-79",
                               TRUE ~ as.character(age)),
         race_group = case_when(hispanic_latino == "Yes" ~ "Hispanic/Latino",
                                race == "White" ~ "Non-Hispanic White",
                                race == "Black or African American" ~ "Non-Hispanic Black",
                                TRUE ~ "Other"),
         memory_rank = stringr::str_to_upper(memory_rank))

elig <- dbReadTable(con, "p2_redcap_eligibility") %>%
  select(record_id, hml_id, not_eligible, hif_eligible_behav, hif_eligible_mri, hif_eligible_bd)

cons <- dbReadTable(con, "p2_redcap_consent_form") %>%
  select(record_id, hml_id, main_consent_date, main_consent_upload)

# Get list of consented participants
consented_ids <- subset(cons, !is.na(main_consent_upload))$hml_id

phq9 <- dbReadTable(con, "p2_redcap_cognitive_function_phq9")

moca <- dbReadTable(con, "p2_redcap_cognitive_function_moca")

naart <- dbReadTable(con, "p2_redcap_cognitive_function_naart")

avlt <- dbReadTable(con, "p2_redcap_cognitive_function_avlt")

biometrics <- dbReadTable(con, "p2_redcap_biometrics")

# dm_checklist <- dbReadTable(con, "p2_redcap_dm_checklist") %>%
  # select(hml_id, dm_checklist_complete)

base <- full_join(phq9, moca, by = c("hml_id", "record_id", "participant_id_parent")) %>%
  full_join(naart, by = c("hml_id", "record_id", "participant_id_parent")) %>%
  full_join(avlt, by = c("hml_id", "record_id", "participant_id_parent")) %>%
  full_join(biometrics, by = c("hml_id", "record_id", "participant_id_parent")) %>%
  # full_join(neuropsych, by = c("hml_id", "record_id", "participant_id_parent")) %>%
  # One of the record_ids is missing a hml_id
  left_join(select(demo, record_id, hml_id) %>% rename(new_hml_id = hml_id), by = "record_id") %>%
  mutate(hml_id = ifelse(is.na(hml_id), new_hml_id, hml_id)) %>%
  select(-new_hml_id)

# surv <- dbReadTable(con, "p2_redcap_survey_info") %>%
#   select(record_id, hml_id, hml_survey_tracker_15_complete)

```

`r if(sum(duplicated(demo$record_id)) > 0){"dm_redcapx_demographic update contains duplicated rows"}`
`r if(sum(duplicated(elig$record_id)) > 0){"dm_redcapx_eligibility update contains duplicated rows"}`
`r if(sum(duplicated(cons$record_id)) > 0){"dm_redcapx_consents update contains duplicated rows"}`
`r if(sum(duplicated(base$record_id)) > 0){"dm_redcapx_baseline_assessment update contains duplicated rows"}`

The following tables summarize data collected in REDCap for PAN Project 2.

Report Date: **`r format(Sys.Date(), "%b %d, %Y")`**

#### Participants with HML IDs

These participants are in the MC Demographics form and have been assigned HML IDs.

```{r recruit summary}

participants_dat <- demo %>%
  left_join(cons, by = "record_id") %>%
  mutate(consent_uploaded = ifelse(!is.na(main_consent_upload), "Consented", "Non-Consented"))

label(participants_dat$sex) <- "Sex"
label(participants_dat$age_group) <- "Age"
label(participants_dat$race_group) <- "Race"
label(participants_dat$consent_uploaded) <- "Consent Uploaded?"

table1(~sex + age_group + race_group + consent_uploaded | area, data = participants_dat)

```

#### Consented Participants

The following table contains only participants whose consent has been uploaded to REDCap. The rest of this report follows these participants.

```{r consented summary}

table1(~sex + age_group + race_group | area, 
       data = filter(participants_dat,  !is.na(main_consent_upload)))

```

```{r nonconsented}

non_consents <- cons %>%
  filter(record_id %in% base$record_id & is.na(main_consent_upload)) %>%
  left_join(select(demo, record_id, area), by = "record_id") %>%
  select(area, record_id, hml_id, main_consent_date, main_consent_upload)

```

<br>

There are `r nrow(non_consents)` participants without uploaded consents, but still have data collected in other REDCap forms. `r nrow(non_consents[!is.na(non_consents$main_consent_date),])` of these have a "Main Consent Date" in the Consent form:

```{r nonconsented table}
if(nrow(non_consents) > 0){
  non_consents %>%
    mutate(across(-c(area, record_id, hml_id), ~ifelse(is.na(.), "Missing", .))) %>%
    kable(col.names = c("Area", "Record ID", "HML ID", "Main Consent Date", "Main Consent Uploaded"),
          align = "c", escape = FALSE) %>%
    column_spec(1:2, bold = TRUE) %>%
    column_spec(1, extra_css = "border-bottom: 1px dashed; background: white;") %>%
    row_spec(0, bold = TRUE) %>%
    collapse_rows(columns = 1) %>%
    kable_classic(full_width = FALSE, lightable_options = "striped", position = "left")
}else{
  print("All participants with REDCap data have uploaded consents.")
}


```

I have removed those participants from the rest of this report.

------

```{r filtering consented participants}

base <- filter(base,
               record_id %in% participants_dat$record_id[participants_dat$consent_uploaded == "Consented"])

elig <- filter(elig,
               record_id %in% participants_dat$record_id[participants_dat$consent_uploaded == "Consented"])

```

### Cognitive Testing and Biometrics

This table summarizes the cognitive testing in Scored Forms and the Biometrics Survey Status from the Experimental Tasks form. Participants without any missing form data are marked as "Non-Missing", if they have at least one missing entry then they are marked as "Missing."

```{r cognitive and biometrics summary}

assess <- base %>%
  # Add in demo data
  full_join(select(demo, record_id, area, hml_id,
                   age_group, race_group, memory_rank, sex), by = c("record_id", "hml_id")) %>%
  # Filter for consented participants
  left_join(cons, by = c("record_id", "hml_id")) %>%
  filter(!is.na(main_consent_upload)) %>%
  # Get postgres update date
  mutate(update_date = format(Sys.Date(), "%y%m%d")) %>%
  # Cognitive Summaries ~~~~~~~~~~~~~~
  mutate(across(phq9_endorse_q9:avlt_trial_a7_raw, ~ifelse(is.na(.), 1, 0))) %>%
  mutate("PHQ-9" = ifelse((phq9_endorse_q9 + phq9_total) > 0, "Missing", "Non-Missing"),
         # 20-Feb-24: Removed the education bonus and mis_total variables since not used in final total:
         MoCA = ifelse(select(., moca_abstraction, starts_with("moca_attention"), moca_delay_recall, starts_with("moca_language_"),
                              moca_visuospatial, moca_naming, moca_orientation, moca_total) %>% rowSums() > 0, "Missing", "Non-Missing"),
         NAART = ifelse(select(., naart_total_correct) %>% rowSums() > 0, "Missing", "Non-Missing"),
         AVLT = ifelse(select(., starts_with("avlt_") & ends_with("_raw")) %>% rowSums() > 0, "Missing", "Non-Missing")) %>%
  # Biometrics Form ~~~~~~~~~~~~~~~~~~
  mutate(across(bp_r_systolic_1:waist_hip_ratio, ~ifelse(is.na(.), 1, 0))) %>%
  mutate("Biometrics Form" = ifelse(select(., bp_r_systolic_1:waist_hip_ratio) %>% rowSums() > 0, "Missing", "Non-Missing")) %>%
  # All table variables
  mutate(across(c(`PHQ-9`, `MoCA`, `NAART`, `AVLT`, 
                  # `Neuropsychology Upload`, 
                  # `PHQ-9 Upload`, 
                  `Biometrics Form`), 
                ~factor(., levels = c("Non-Missing", "Missing")))) %>%
  # Get overall missingness:
  mutate(Overall = factor(ifelse(`PHQ-9` == "Non-Missing" & MoCA == "Non-Missing" & NAART == "Non-Missing" & 
                                   AVLT == "Non-Missing" & 
                                   # `Neuropsychology Upload` == "Non-Missing" & 
                                   # `PHQ-9 Upload` == "Non-Missing" & 
                                   `Biometrics Form` == "Non-Missing", "Complete Neuropsych and Biometrics", 
                                 "Incomplete Neuropsych and Biometrics"),
                          levels = c("Complete Neuropsych and Biometrics", "Incomplete Neuropsych and Biometrics"))) %>%
  select(record_id, hml_id = hml_id, area, update_date, age_group, race_group, memory_rank, sex, main_consent_date, ends_with("_complete"),
         main_consent_upload, `PHQ-9`, `MoCA`, `NAART`, `AVLT`, 
         # `Neuropsychology Upload`, 
         # `PHQ-9 Upload`, 
         `Biometrics Form`,
         Overall) %>%
  # Add in whether participant is eligible for study
  left_join(elig, by = c("record_id", "hml_id" = "hml_id")) %>%
  mutate(not_eligible = case_when(not_eligible == 0 ~"Not Eligible", 
                                  !is.na(hif_eligible_behav) ~ "Eligible"))

table1(~`PHQ-9` + MoCA + NAART + AVLT  +  `Biometrics Form` + Overall| area, 
       data = assess)
  
```

```{r list of hmls with complete data, eval = FALSE}

hml_complete <- assess %>%
  filter(`PHQ-9` == "Non-Missing", MoCA == "Non-Missing", NAART == "Non-Missing", AVLT == "Non-Missing", `Neuropsychology Upload` == "Non-Missing", `PHQ-9 Upload` == "Non-Missing", `Biometrics Form` == "Non-Missing") %>%
  select(area, hml_id)

atlanta_ignore_moca <- assess %>%
  filter(area %in% "Atlanta", `PHQ-9` == "Non-Missing", NAART == "Non-Missing", AVLT == "Non-Missing", `Neuropsychology Upload` == "Non-Missing", `PHQ-9 Upload` == "Non-Missing", `Biometrics Form` == "Non-Missing") %>%
  select(area, hml_id) %>%
  slice_sample(n = 20)

combined_hml_complete <- hml_complete %>%
  rbind(atlanta_ignore_moca) %>%
  arrange(desc(area), hml_id)

# write.csv(combined_hml_complete, file = "complete_data_hml_ids_240124.csv")

```

--------

### Other Core and Project Data

```{r frailty_data}

# Analyzed data
analyzed_frailty <- dbReadTable(con, "coref_najafi_frailty") %>%
  mutate(complete_frailty = "Complete") %>%
  select(hml_id, complete_frailty) %>% distinct()

```

```{r garmin_data}

analyzed_garmin <- dbReadTable(con, "coref_najafi_garmin") %>%
  mutate(complete_garmin = ifelse(days_of_recording > 0, "Complete", "Incomplete")) %>%
  select(hml_id, complete_garmin) %>% distinct()

```

```{r language_tasks}

# Analyzed data
analyzed_lang <- dbReadTable(con, "coref_mehl_egemaps") %>%
  # Each ID has two rows so need to count how many people have both rows in data
  group_by(hml_id) %>% mutate(n = n()) %>% ungroup() %>% filter(n == 2) %>%
  select(hml_id) %>% distinct() %>%
  mutate(complete_lang = "Complete")

```

```{r sonora_quest}

# Analyzed data
analyzed_sonora_quest <- dbReadTable(con, "p2_sonora_quest_labwork") %>%
  select(hml_id) %>% distinct() %>% mutate(complete_sonora_quest = "Complete")

```

```{r carotid_ultrasound}

analyzed_carotid_ultrasound <- dbReadTable(con, "p2_carotid_ultrasound") %>%
  mutate(complete_carotid_ultrasound = ifelse(is.na(notes) | notes != "Unable to open images in XNAT", "Complete", "Incomplete")) %>%
  select(hml_id, complete_carotid_ultrasound) %>% distinct()

```

```{r apoe}
coreg_huentelman_apoe4 <- dbReadTable(con, "coreg_huentelman_apoe4") %>%
  mutate(complete_apoe = "Complete") %>%
  select(hml_id, complete_apoe)
```

```{r ad_neurofilament_assay}
coreg_ad_plasma_biomarkers <- dbReadTable(con, "coreg_ad_plasma_biomarkers") %>%
  mutate(complete_ad_biomarkers = "Complete") %>%
  select(hml_id, complete_ad_biomarkers)
```

```{r cytokine_data}
coreg_doyle_cytokine <- dbReadTable(con, "coreg_doyle_cytokine") %>%
  select(hml_id, Eotaxin, Fgf2, IFNg, IL_10, IL_12p40, IL_13, IL_15, IL_17a, IL_1a, IL_1b, IL_2, IL_4, IL_5, IL_6, IL_7, IL_8, IP_10, MCP_1, MDC, MIP_1a, MIP_1b, TNFa, TNFb, VEGFa) %>%
  # If ID has data in column assign TRUE, else FALSE
  mutate(across(-hml_id, ~ifelse(!is.na(.), TRUE, FALSE))) %>%
  # Sum across non-ID columns
  mutate(completed = select(., -hml_id) %>% rowSums(na.rm = TRUE),
         complete_cytokine = ifelse(completed == 24, "Complete", "Incomplete")) %>%
  select(hml_id, complete_cytokine)
```

```{r epigenetic}

coreg_huentelman_epigenetic <- dbReadTable(con, "coreg_huentelman_epigenetic") %>%
  mutate(complete_epi = "Complete") %>%
  select(hml_id, complete_epi)

```

```{r prs}

coreg_huentelman_prs <- dbReadTable(con, "coreg_huentelman_prs") %>%
  mutate(complete_prs = "Complete") %>%
  select(hml_id, complete_prs)

```

```{r cortisol}
coref_runyon_cortisol <- dbReadTable(con, "coref_runyon_cortisol") %>%
  mutate(complete_cortisol = ifelse(!is.na(cortisol), "Complete", "Incomplete")) %>%
  select(hml_id, complete_cortisol)
```

```{r mri}

coree_chen_anat_measures <- dbReadTable(con, "coree_chen_anat_measures") %>%
  mutate(complete_anat = "Complete") %>%
  select(hml_id, complete_anat)

coree_chen_asl_measures <- dbReadTable(con, "coree_chen_asl_measures") %>%
  mutate(complete_asl = "Complete") %>%
  select(hml_id, complete_asl)

coree_chen_dwi_measures <- dbReadTable(con, "coree_chen_dwi_measures") %>%
  mutate(complete_dwi = "Complete") %>%
  select(hml_id, complete_dwi)

coree_chen_rsfmri <- dbReadTable(con, "coree_chen_rsfmri_yeobuckner131_interconn_corticocereb") %>%
  mutate(complete_rsfmri = "Complete") %>%
  select(hml_id, complete_rsfmri)

```

```{r combining_table_other_data}

# Get list of consented participants
assay_table_dat <- data.frame(hml_id = consented_ids) %>%
  # P2
  left_join(analyzed_carotid_ultrasound, by = "hml_id") %>%
  left_join(analyzed_sonora_quest, by = "hml_id") %>%
  # left_join(analyzed_surveys, by = "hml_id") %>%
  # left_join(analyzed_survey_info, by = "hml_id") %>%
  # Core E
  left_join(coree_chen_anat_measures, by = "hml_id") %>%
  left_join(coree_chen_asl_measures, by = "hml_id") %>%
  left_join(coree_chen_dwi_measures, by = "hml_id") %>%
  left_join(coree_chen_rsfmri, by = "hml_id") %>%
  # Core F
  left_join(coref_runyon_cortisol, by = "hml_id") %>%
  left_join(analyzed_frailty, by = "hml_id") %>%
  left_join(analyzed_garmin, by = "hml_id") %>%
  left_join(analyzed_lang, by = "hml_id") %>%
  # Core G
  left_join(coreg_ad_plasma_biomarkers, by = "hml_id") %>%
  left_join(coreg_huentelman_apoe4, by = "hml_id") %>%
  left_join(coreg_doyle_cytokine, by = "hml_id") %>%
  left_join(coreg_huentelman_epigenetic, by = "hml_id") %>%
  left_join(coreg_huentelman_prs, by = "hml_id") %>%
  # # Calculate total missing or incomplete for each person
  mutate_all(~ifelse(is.na(.), "Incomplete", .)) %>%
  # mutate(across(-hml_id, ~ifelse(. == "Incomplete", TRUE, FALSE))) %>%
  # mutate(all_data = select(., -hml_id) %>% rowSums(na.rm = TRUE),
  #        all_data = ifelse(all_data == 0, "Complete", "Incomplete")) %>%
  # mutate(across(-c(hml_id, all_data), ~ifelse(., "Incomplete", "Complete"))) %>%
  left_join(select(demo, hml_id, site = area), by = "hml_id") %>%
  filter(!is.na(site))

# Wanted to use table1, but need to figure out notes
# rndr <- function(x, ...) {
#     y <- render.default(x, ...)
#     if (is.factor(x)) y[2] else y
# }
# 
# table1(~. | site, assay_table_dat[-1], render = rndr)

```

```{r}
assay_table_dat_calc <- assay_table_dat %>%
  rbind(assay_table_dat %>% mutate(site = "Overall")) %>%
  # filter(hml_id %in% consented_ids_sub) %>%
  mutate(across(-c(hml_id, site), ~ifelse(. == "Complete", 1, 0))) %>%
  mutate(`All Data` = select(., -c(hml_id, site)) %>% rowSums(na.rm = TRUE),
         # Assign 1 if participant has all data available
         `All Data` = ifelse(`All Data` == ncol(select(., -c(hml_id, site))), 1, 0),
         site = factor(site,
                       levels = c("Atlanta", "Baltimore", "Miami", "Tucson", "Overall"))) %>%
  # Group by site
  select(-hml_id) %>% group_by(site) %>%
  # Get number for each group and sum completed data
  mutate(num = n()) %>% 
  group_by(num, .add = T) %>% summarise(across(everything(), sum), .groups = "keep") %>%
  # Add percentages
  mutate(across(everything(), ~paste0(., " (", round(./num*100,1), "%)")),
         site = paste0(site, " (n=", num, ")")) %>%
  rename(`Carotid Ultrasound` = complete_carotid_ultrasound,
         `Frailty` = complete_frailty,
         `Garmin` = complete_garmin,
         `Anatomical MRI` = complete_anat,
         `ASL MRI` = complete_asl,
         `Diffusion-weighted MRI` = complete_dwi,
         `Resting-state fMRI` = complete_rsfmri,
         `Language Tasks` = complete_lang,
         `Sonora Quest` = complete_sonora_quest,
         # `Surveys (Complete Data)` = complete_surveys,
         # `Survey REDCap Tracker*` = complete_survey_info,
         `Cortisol` = complete_cortisol,
         `AD (Neurofilament Assay)` = complete_ad_biomarkers,
         `APOE4` = complete_apoe,
         `Cytokine` = complete_cytokine,
         `Epigenetic Clock` = complete_epi,
         `Polygenic Risk Score` = complete_prs) %>%
  ungroup() %>% select(-num) %>%
  pivot_longer(cols = -site,
               names_to = "Data") %>%
  pivot_wider(id_cols = Data,
              names_from = site, values_from = value)

kable(cbind(data.frame(Source = c(rep("P2", 2), rep("Core E", 4), rep("Core F", 4), rep("Core G", 5), "All Data")),
            assay_table_dat_calc)) %>%
  kable_classic() %>%
  row_spec(0, bold = T) %>%
  column_spec(1, bold = T) %>%
  collapse_rows(1) %>%
  row_spec(c(2, 6, 10, 15), extra_css = "border-bottom: 1px solid")

```

------------

### MindCrowd Expanded Cognitive Function Tasks And Surveys

The following summaries are for Health Minds for Life (P2) participants only.

#### MindCrowd Cognitive Function Tasks

```{r mindcrowd_tasks}

scores_flk <- dbGetQuery(con, "SELECT * FROM scores_flk WHERE hml_id is not NULL") %>%
  filter(FLK_first_attempt) %>%
  mutate(flk_complete = ifelse(FLK_qc_code == "QC passed", "Complete", "Incomplete")) %>%
  select(hml_id, flk_complete)

scores_kt <- dbGetQuery(con, "SELECT * FROM scores_kt WHERE hml_id is not NULL") %>%
  filter(KT_first_attempt) %>%
  mutate(kt_complete = ifelse(KT_qc_code == "QC passed", "Complete", "Incomplete")) %>%
  select(hml_id, kt_complete)

scores_ln <- dbGetQuery(con, "SELECT * FROM scores_ln WHERE hml_id is not NULL") %>%
  filter(LN_first_attempt) %>%
  mutate(ln_complete = ifelse(LN_qc_code == "QC passed", "Complete", "Incomplete")) %>%
  select(hml_id, ln_complete)

scores_scrt <- dbGetQuery(con, "SELECT * FROM scores_scrt WHERE hml_id is not NULL") %>%
  filter(SCRT_first_attempt) %>%
  mutate(scrt_complete = ifelse(SCRT_qc_code == "QC passed", "Complete", "Incomplete")) %>%
  select(hml_id, scrt_complete)

scores_wp <- dbGetQuery(con, "SELECT * FROM scores_wp WHERE hml_id is not NULL") %>%
  filter(WP_first_attempt) %>%
  mutate(wp_complete = ifelse(WP_qc_code == "QC passed", "Complete", "Incomplete")) %>%
  select(hml_id, wp_complete)

cog_scores <- data.frame(hml_id = consented_ids) %>%
  left_join(scores_flk, by = "hml_id") %>%
  left_join(scores_kt, by = "hml_id") %>%
  left_join(scores_ln, by = "hml_id") %>%
  left_join(scores_scrt, by = "hml_id") %>%
  left_join(scores_wp, by = "hml_id") %>%
  mutate(across(-hml_id, ~ifelse(is.na(.), "Incomplete", "Complete"))) %>%
  mutate(scores_complete = ifelse(flk_complete == "Complete" & kt_complete == "Complete" &
                                  ln_complete == "Complete" & scrt_complete == "Complete" &
                                  wp_complete == "Complete", "Complete", "Incomplete")) %>%
  distinct() %>%
  left_join(select(demo, hml_id, site = area), by = "hml_id") %>%
  filter(!is.na(site))

cog_scores_table <- cog_scores %>%
  rbind(cog_scores %>% mutate(site = "Overall")) %>%
  mutate(across(-c(hml_id, site), ~ifelse(. == "Complete", 1, 0))) %>%
  mutate(site = factor(site,
                       levels = c("Atlanta", "Baltimore", "Miami", "Tucson", "Overall"))) %>%
  # Group by site
  select(-hml_id) %>% group_by(site) %>%
  # Get number for each group and sum completed data
  mutate(num = n()) %>% 
  group_by(num, .add = T) %>% summarise(across(everything(), sum), .groups = "keep") %>%
  # Add percentages
  mutate(across(everything(), ~paste0(., " (", round(./num*100,1), "%)")),
         site = paste0(site, " (n=", num, ")")) %>%
  rename("Flanker" = "flk_complete",
         "Keep Track" = "kt_complete",
         "Letter Number" = "ln_complete",
         "Simple and Choice Reaction Time" = "scrt_complete",
         "Word Pairs" = "wp_complete",
         "All Cognitive Function Tasks" = "scores_complete") %>%
  ungroup() %>% select(-num) %>%
  pivot_longer(cols = -site,
               names_to = "Data") %>%
  pivot_wider(id_cols = Data,
              names_from = site, values_from = value)
  
kable(cog_scores_table) %>%
  kable_classic() %>%
  row_spec(0, bold = T) %>%
  column_spec(1, bold = T) %>%
  row_spec(c(5), extra_css = "border-bottom: 1px solid")

# label(cog_scores$flk_complete) <- "Flanker"
# label(cog_scores$kt_complete) <- "Keep Track"
# label(cog_scores$ln_complete) <- "Letter Number"
# label(cog_scores$scrt_complete) <- "Simple and Choice Reaction Time"
# label(cog_scores$wp_complete) <- "Word Pairs"
# label(cog_scores$scores_complete) <- "All Cognitive Tasks"
# 
# table1(~ . | site, cog_scores[-1])

```

#### Surveys

```{r surveys}

# Get survey data --------------
survey_names <- c("adl", "anxiety", "brain_disease", "covid", "diet", "fhad", 
           "health_medical", "perceived_stress", "qpar", "ses", "sleep",
           "social_stressor", "social_support", "subjective_english", "swls")

raw_survey_list <- lapply(survey_names, 
                          function(x){
                            if(x == "fhad"){
                              new_x <- dbReadTable(con, paste0("p2_redcap_", "family_history"))
                            }else{
                              new_x <- dbReadTable(con, paste0("p2_redcap_", x))
                            }
                            
                            x <- case_when(x == "health_medical" ~ "healthmed",
                                           x == "perceived_stress" ~ "stress",
                                           x == "social_stressor" ~ "socstress",
                                           x == "social_support" ~ "socsupp",
                                           x == "subjective_english" ~ "subeng",
                                           TRUE ~ x)
                            
                            new_x %>%
                              select(hml_id, all_of(paste0(x, "_complete")))
                            
                          })
names(raw_survey_list) <- survey_names

completed_surveys <- reduce(raw_survey_list, full_join, by = "hml_id") %>%
  mutate(across(-hml_id, ~ifelse(is.na(.), 0, .))) %>%
  # Switch value for incomplete surveys to 1 and complete to 0 so I can sum up number of incomplete surveys
  mutate(across(-hml_id, ~ifelse(. == 1, 0, 1))) %>%
  # Sum up incomplete surveys
  mutate(incomplete_total = select(., ends_with("_complete")) %>% rowSums(na.rm = TRUE)) %>%
  mutate(surveys_complete = ifelse(incomplete_total > 0, "Incomplete", "Complete")) %>%
  # Change 1/0s back to incomplete/complete
  mutate(across(-c(hml_id, incomplete_total, surveys_complete), ~ifelse(. == 0, "Complete", "Incomplete"))) %>%
  select(-c(incomplete_total))

# Get REDCap coordinator check data
analyzed_survey_info <- dbReadTable(con, "p2_redcap_survey_info") %>%
  mutate(complete_survey_info = ifelse(hml_survey_tracker_complete == 2, "Complete", "Incomplete")) %>%
  select(hml_id, complete_survey_info)

```

```{r compile surveys}

survey_dat <- data.frame(hml_id = consented_ids) %>%
  left_join(completed_surveys, by = "hml_id") %>%
  left_join(analyzed_survey_info, by = "hml_id") %>%
  left_join(select(demo, hml_id, site = area), by = "hml_id") %>%
  mutate(across(-hml_id, ~ifelse(is.na(.), "Incomplete", .)))

survey_dat_calc <- survey_dat %>%
  rbind(survey_dat %>% mutate(site = "Overall")) %>%
  mutate(across(-c(hml_id, site), ~ifelse(. == "Complete", 1, 0))) %>%
  mutate(site = factor(site,
                       levels = c("Atlanta", "Baltimore", "Miami", "Tucson", "Overall"))) %>%
  # Group by site
  select(-hml_id) %>% group_by(site) %>%
  # Get number for each group and sum completed data
  mutate(num = n()) %>% 
  group_by(num, .add = T) %>% summarise(across(everything(), sum), .groups = "keep") %>%
  # Add percentages
  mutate(across(everything(), ~paste0(., " (", round(./num*100,1), "%)")),
         site = paste0(site, " (n=", num, ")")) %>%
  rename("ADL" = "adl_complete",
         "Anxiety" = "anxiety_complete",
         "Brain Disease" = "brain_disease_complete",
         "Covid-19" = "covid_complete",
         "Diet" = "diet_complete",
         "FHAD" = "fhad_complete",
         "Health & Medical" = "healthmed_complete",
         "Stress" = "stress_complete",
         "QPAR" = "qpar_complete",
         "SES" = "ses_complete",
         "Sleep" = "sleep_complete",
         "Social Stressors" = "socstress_complete",
         "Social Support" = "socsupp_complete",
         "Subjective English" = "subeng_complete",
         "SWLS" = "swls_complete",
         "All Surveys" = "surveys_complete",
         "REDCap Survey Tracker*" = "complete_survey_info") %>%
  ungroup() %>% select(-num) %>%
  pivot_longer(cols = -site,
               names_to = "Data") %>%
  pivot_wider(id_cols = Data,
              names_from = site, values_from = value)

kable(survey_dat_calc) %>%
  kable_classic() %>%
  row_spec(0, bold = T) %>%
  column_spec(1, bold = T) %>%
  row_spec(c(15, 16), extra_css = "border-bottom: 1px solid") %>%
  footnote(symbol = c( "Verified in REDCap, but some entries contain missing data."))

```

--------------------------

#### Participants marked as ineligible or dropped from the study:

The following were ineligible for behavioral testing:

- `r elig[which(elig$hif_eligible_behav == "No"), "hml_id"]`

The following were ineligible for MRI scanning:

- `r elig[which(elig$hif_eligible_mri == "No"), "hml_id"]`

The following were ineligible for blood draw:

- `r elig[which(elig$hif_eligible_bd == "No"), "hml_id"]`

IDs dropped from the study:

```{r eligibility}

elig %>% 
  left_join(select(demo, record_id, area), by = "record_id") %>%
  filter(not_eligible == 0) %>%
  select(hml_id, record_id, area, hif_eligible_behav, hif_eligible_mri, hif_eligible_bd) %>%
  kable(col.names = c("HML ID", "REDCap ID", "Area", "Eligible for Behavior?", "Eligible for MRI?", "Eligible for Blood Draw?")) %>% 
  kable_classic(full_width = FALSE)

```

<!-- #### Data Flowchart -->

```{r flowchart data setup, eval = FALSE}

flow_dat <- demo %>%
  # Add consent data
  left_join(cons, by = c("record_id", "hml_id")) %>%
  mutate(consent_uploaded = ifelse(!is.na(main_consent_upload), "Consented", "Non-Consented")) %>%
  # Add eligibility data
  left_join(elig, by = c("record_id", "hml_id")) %>%
  # Add cognitive and biometrics data
  left_join(base, by = c("record_id", "hml_id")) %>%
  # Cognitive Summaries ~~~~~~~~~~~~~~
  mutate(across(phq9_endorse_q9:avlt_trial_a7_raw, ~ifelse(is.na(.), 1, 0))) %>%
  mutate("PHQ-9" = ifelse((phq9_endorse_q9 + phq9_total) > 0, "Missing", "Non-Missing"),
         # 20-Feb-24: Removed the education bonus and mis_total variables since not used in final total:
         MoCA = ifelse(select(., moca_abstraction, starts_with("moca_attention"), moca_delay_recall, starts_with("moca_language_"),
                              moca_visuospatial, moca_naming, moca_orientation, moca_total) %>% rowSums() > 0, "Missing", "Non-Missing"),
         NAART = ifelse(select(., naart_total_correct) %>% rowSums() > 0, "Missing", "Non-Missing"),
         AVLT = ifelse(select(., starts_with("avlt_") & ends_with("_raw")) %>% rowSums() > 0, "Missing", "Non-Missing")) %>%
  # Biometrics Form ~~~~~~~~~~~~~~~~~~
  mutate(across(bp_r_systolic_1:waist_hip_ratio, ~ifelse(is.na(.), 1, 0))) %>%
  mutate("Biometrics Form" = ifelse(select(., bp_r_systolic_1:waist_hip_ratio) %>% rowSums() > 0, "Missing", "Non-Missing")) %>%
  # All table variables
  mutate(across(c(`PHQ-9`, `MoCA`, `NAART`, `AVLT`, 
                  # `Neuropsychology Upload`, 
                  # `PHQ-9 Upload`, `Biometrics Form`
                  ), 
                ~factor(., levels = c("Non-Missing", "Missing")))) %>%
  # Get overall missingness:
  mutate(Overall = factor(ifelse(`PHQ-9` == "Non-Missing" & MoCA == "Non-Missing" & NAART == "Non-Missing" & 
                                   AVLT == "Non-Missing" & 
                                   # `Neuropsychology Upload` == "Non-Missing" & 
                                   # `PHQ-9 Upload` == "Non-Missing" & 
                                   `Biometrics Form` == "Non-Missing", "Complete Neuropsych and Biometrics", 
                                 "Incomplete Neuropsych and Biometrics"),
                          levels = c("Complete Neuropsych and Biometrics", "Incomplete Neuropsych and Biometrics")))

```

```{r flowchart, eval = FALSE}

# Flowchart outline

# Assign exclusion (remove) variables and reasoning if excluded; should be NA if participant is not excluded
# Assign inclusion variables that entered the next phase

# all hml_ids - record_id

# remove non-consented - is.na(main_consent_upload)
flow_dat$non_consented <- ifelse(!is.na(flow_dat$main_consent_upload), NA, "Not Consented")
# consented
flow_dat$consented_group <- ifelse(!is.na(flow_dat$non_consented), NA, flow_dat$record_id)

# remove non-eligble
flow_dat$non_eligible <- case_when(is.na(flow_dat$not_eligible) ~ NA,
                                           flow_dat$hif_eligible_behav == "No" ~ "Ineligble for Behavioral Testing",
                                           flow_dat$hif_eligible_mri == "No" ~ "Ineligble for MRI",
                                           flow_dat$hif_eligible_bd == "No" ~ "Ineligble for Blood Draw")
# eligble participants
flow_dat$eligible <- ifelse(!is.na(flow_dat$non_consented) | !is.na(flow_dat$not_eligible), NA, flow_dat$record_id)

# remove missing cognitive/biometrics data
flow_dat$missing_data <- case_when(flow_dat$`PHQ-9` == "Missing" ~ "PHQ-9",
                                          flow_dat$MoCA == "Missing" ~ "MoCA",
                                          flow_dat$NAART == "Missing" ~ "NAART",
                                          flow_dat$AVLT == "Missing" ~ "AVLT",
                                          # flow_dat$`Neuropsychology Upload` == "Missing" ~ "Neuropsych Upload",
                                          flow_dat$`Biometrics Form` == "Missing"~ "Biometrics",
                                          TRUE ~ NA)

# non-missing participants
flow_dat$non_missing <- ifelse((!is.na(flow_dat$non_consented) | !is.na(flow_dat$not_eligible) |
                                         !is.na(flow_dat$missing_data)), NA, flow_dat$record_id)

consort_plot(flow_dat,
             orders = c(record_id = "Screened",
                        # area = "Site",
                        non_consented = "Not Consented",
                        consented_group = "Consented",
                        non_eligible = "Ineligible",
                        eligible = "Eligible",
                        missing_data = "Missing Cognitive/Biometrics Data",
                        non_missing = "Non-Missing Data"),
             side_box = c("non_consented", "non_eligible", "missing_data")
             # ,allocation = "area"
             )

```

--------------------------